---
title: "p8105_hw3_cw3555"
author: "Eunice Wang"
date: '`2023-10-14`'
output: github_document

---

```{r}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 1

#### Read in the data

```{r}
data("instacart")

instacart = 
  instacart |> 
  as_tibble()
```

#### Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row representing a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart |> select(product_id) |> distinct() |> count()` products found in `r instacart |> select(user_id, order_id) |> distinct() |> count()` orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```

### Problem 2

#### Read in the data

```{r}
data("brfss_smart2010")
```

#### Data cleaning

```{r}
brfss_smart2010_df = 
  brfss_smart2010 |> 
  janitor::clean_names() |> 
  filter(
    topic=="Overall Health",
    response %in% c("Excellent","Very good","Good","Fair","Poor")
 ) |> 
mutate(
  response = factor(response, levels= c("Poor", "Fair", "Good", "Very good", "Excellent"), ordered= TRUE)
 )
```



#### Using data to answer the following questions

* Q: In 2002, which states were observed at 7 or more locations? 

```{r}
brfss_smart2010_df |> 
  filter(year==2002) |> 
  group_by(year, locationabbr) |> 
  summarize(distinct_locations = n_distinct(locationdesc)) |> 
  filter(distinct_locations>=7)|> 
  knitr::kable()
```

* A: In 2002, total 6 states are observed at 7 or more locations: CT, FL, MA, NC, NJ, PA.

* Q: What about in 2010?

```{r}
brfss_smart2010_df |> 
  filter(year==2010) |> 
  group_by(year, locationabbr) |> 
  summarize(distinct_locations = n_distinct(locationdesc)) |> 
  filter(distinct_locations>=7) |> 
  knitr::kable()
```

* A: In 2002, total 14 states are observed at 7 or more locations: CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, WA.

* Q: Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
brfss_excellent_repsonse_df=
  brfss_smart2010_df |> 
  filter(response=="Excellent") |> 
  group_by(year, locationabbr) |> 
  summarize(mean_value = mean(data_value, na.rm =TRUE))

brfss_excellent_repsonse_df |> 
  ggplot(aes(x=year, y=mean_value, color= locationabbr))+
  geom_line(alpha=0.7)+
labs(
    title = "Average Value of 'Excellent' Health Response Over Time Within a State",
    x= "Year(2002-2010)",
    y= "Average Value of Overall Health Reponse",
    caption = "Data of 'Excellent Reponses' from BRFSS data"
  )+
   theme_classic()
```

A: This above “spaghetti” plot shows the average value of "excellent" health response over time within a state. Each line with different colors represent different states. Most states have a volatile pattern from 2002-2010, containing both increasing and decreasing trends. Though they all perceive their overall health as excellent, there are times that they consider it to be improved and worsened.

* Q: Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State

```{r}
brfss_2006_2010_NY_df =
  brfss_smart2010_df |> 
  filter(year%in% c("2006", "2010"), locationabbr == "NY") 

brfss_2006_2010_NY_df |> 
  ggplot(aes(x=response, y= data_value))+
  geom_boxplot()+
  facet_grid(.~ year)+
  labs(
    title = "Distirbution of Value of Health Response for 2006 and 2010 in NY",
    x= "Response(Poor-Excellent)",
    y= "Value of Overall Health Reponse",
    caption = "Data of 'Year 2006 & 2010 in NY' from BRFSS data"
  ) 
```

* A: This two-panel box plot shows the distribution of value of health response ((“Poor” to “Excellent”) for 2006 and 2010 in NY. Comparing the two panel, it can be inferred that there are some improvements in one's perception of their over health comparing 2006 and 2010. The median value of "Very good" response increased and the 25th and 75th interquartile also increased. There are also slight increases in median value of "Fair"and "Good" response.

### Problem 3

#### Load and tidy the data

```{r}
nhanes_covar_df =
  read.csv("problem_3_data/nhanes_covar.csv", skip=4) |> 
  janitor::clean_names() |> 
  mutate(
    sex= case_match(
      sex,
      1~ "Male",
      2~ "Female"
    ),
    education = case_match(
      education,
      1 ~ "Less than high school",
      2 ~ "High school equivalent",
      3 ~ "More than high school"
    ), 
    sex= as.factor(sex),
    education=as.factor(education)
  ) |> 
  filter(age>=21) |> 
  drop_na()
  
nhanes_accel_df =
  read.csv("problem_3_data/nhanes_accel.csv") |> 
  janitor::clean_names() |>
  pivot_longer(
    cols = starts_with("min"),
    names_to = "minute",
    values_to = "activity",
    names_prefix = "min"
  ) |> 
  mutate(minute=as.numeric(minute))
```


#### Merge the data

```{r}
nhances_df= 
  left_join(nhanes_covar_df,nhanes_accel_df, by="seqn")
```

#### Produce a reader-friendly table for the number of men and women in each education category

```{r}
  nhances_df |> 
  group_by(sex,education) |> 
  summarize(n_obs=n())|> 
  pivot_wider(
    names_from = sex,
    values_from = n_obs
  ) |> 
  knitr::kable()
```

The table shows that there are more men who are in the high school equivalent category than women , and more women who are in the more than high school(highest education group among the three) and less than high school(lowest education group among the three) ) categories than men.

#### Create a visualization of the age distributions for men and women in each education category.
```{r}
nhances_df |> 
  ggplot(aes(x=education,y=age, fill=sex)) +
  geom_boxplot() +
  labs(title = "Age Distributions for Men and Women in Each Education Category ",
       x= "Education Category",
       y= "Age",
       caption = "Data from merged dataset 'nhances_df'"
       )+
       theme_classic()
```

The plot shows the age distributions for men and women in three different education categories. Age of those in more than high school category is younger than those in less than high school category and high school equivalent category. Age of female in high school equivalent category is less than age of female in less than high school category, while the distrbution of age of men in both categories are similar.

#### Aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis).

```{r}
aggregated_df= 
  nhances_df |> 
  group_by(seqn,sex,age,education) |> 
  summarize(total_activities=sum(activity))

aggregated_df |> 
  ggplot(aes(x=age, y=total_activities,color=sex))+
  geom_point(alpha=0.7)+
  geom_smooth(se = FALSE) +
  facet_grid(.~education)+
  labs(title = "Distirbutrion of Total Activities Against Age for Men and Women in Each Education Category",
       x= "Age",
       y= "Total Activities",
       caption = "Data from aggregated dataset 'aggregated_df'"
       )
```

The panel plot shows the relationship between total activities and age for men and women in different education categories by separating three sub panels. From the smooth line, we can see that there is a negative relationship between total activities and age. As age increases, total activities decrease. In each sub panels, we can see that women generally have more activities than men in all ages excepted when in less than high school category, men who are have more than 40 have more activities than women more than 40. People in more than high school category tends to have a more gentle decreasing trend of activities as age increases compared with the other two categories.

#### Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. 

```{r}
nhances_df |> 
  ggplot(aes(x=minute, y=activity, color=sex))+
  geom_line(alpha=0.3)+
  geom_smooth(se = FALSE)+
  facet_grid(.~education)+
  labs(title = "24-hour Activity Time Courses for Men and Women in Each Education Category",
       x= "Time in Mins of a Day",
       y= "Activity",
       caption = "Data from merged dataset 'nhances_df'"
       )
```

The three-panel plot shows the 24-hour activity time courses(how activities change over time in a day) for men and women in three education levels. From the smooth line in each panel, we can see people tend to have most activities during 500-1000 mins in a day and least activities during 0-250 mins of a day. From the three panels, we can see that the two smooth lines for each gender are close to each other in high school equivalent and less than high school categories, which indicates that men and women have similar trends of activities in a day. However, for the more than high school category, women tends to have more activities in the middle of a day than men. 


